---
title: "CevalSumary"
author: "Xinlong Yang"
date: "2022/11/4"
output: 
  html_document: 
    toc: yes
---

## Setup

```{r setup, results = "hide",message = FALSE,warning = FALSE}
library(tidyverse) #the core packages
library(rstatix)  #in this package t_test can use pipe
library(berryFunctions) # add null rows to a data frame
```

## Import data from .rds

Use **read_rds()** to import data into the session.

```{r,importdata,results = "hide"}
bayer <- read_rds("/home/xinlong.yang/Projects/Ceval/Datasets/bayer.rds")
ceval <- read_rds("/home/xinlong.yang/Projects/Ceval/Datasets/ceval.rds")
```

If you want to import the sas froamt data,must use function: **read_sas** in the package:**haven**.   

```{r,eval = FALSE}
data <- read_sas(file="")
```

## Tidy data

Before analyzing data,Tiding and manipulating are the core phase which make process smoothly.  

```{r}
# filter the part you need
vstall <- bayer %>% filter(ANATYPE=="原始数据") %>% 
  # due to analyze the result of different visit,
  #     adding a variable about visit is nessesary
  cbind(visitn=1) 
```

Codes above show how to filter rows,and insert new variables to a data frame.   
With pipe **%>%**,can omit the input data argument in most functions.

Below will show how to modify values of data frame.   
According rule about judging visit,modifying visit values.

```{r}
vst <- vstall
vst$visitn[vst$CGMTM <= 1440] <- 2  # 1 day
vst$visitn[vst$CGMTM <= 7200 & vst$CGMTM > 1440] <- 3  # 5 day
vst$visitn[vst$CGMTM <= 14400 & vst$CGMTM > 7200] <- 4 # 10 day
vst$visitn[vst$CGMTM <= 21600 & vst$CGMTM > 14400] <- 5  # 15 day
```

But it's a little complex.How to simplify?

```{r}
vst$visitn <- as.integer(cut(vst$CGMTM, breaks=c(0,1440,7200,14400,21600),labels=c(2,3,4,5)))+1
```

Function cut() can be used to simplify the assignment procedure.   Why modifying with adding 1?   

as the output below,cut() returns the factor level defined in argument:labels=c().    
When converting to int, the result will be turned to the simple integer codes related to the sequence of the factors.   
So use another vector to adjust the visit number.

```{r}
test <- c(1320,3000)
cut(test, breaks=c(0,1440,7200,14400,21600),labels=c(2,3,4,5))
as.integer(cut(test, breaks=c(0,1440,7200,14400,21600),labels=c(2,3,4,5)))
as.integer(cut(test, breaks=c(0,1440,7200,14400,21600),labels=c(2,3,4,5)))+1
```

Then assign the stratified variable.

```{r}
st <- rbind(
  vstall,
  vst
)

# index stra fro YSI or CGM
st2 <- rbind(
  mutate(st,stvar=st$YSIRES,stra=1),
  mutate(st,stvar=st$CGMRES,stra=2)
)
```

## Perform analysis

```{r,message = FALSE}
# grouped summarise
mean <- st2 %>% 
  group_by(visitn,stra,CGMLOCN) %>% 
  summarise(
    cnt=format(n(),digits = 0, nsmall = 0),
    mean = format(mean(stvar, na.rm = TRUE),digits = 2, nsmall = 2),
    median=format(median(stvar, na.rm = TRUE),digits = 2, nsmall = 2),
    std=format(sd(stvar, na.rm = TRUE),digits = 3, nsmall = 3),
    min=format(min(stvar),digits = 1, nsmall = 1),
    max=format(max(stvar),digits = 1, nsmall = 1),
    q1=format(quantile(stvar,probs=c(0.25),na.rm=TRUE,type=2),digits = 2, nsmall = 2),
    q3=format(quantile(stvar,probs=c(0.75),na.rm=TRUE,type=2),digits = 2, nsmall = 2),
    
    mnstd=paste(c(mean,'(',std,")"),sep = "",collapse = ""),
    q13=paste(c(q1,',',q3),sep = "",collapse = ""),
    ext=paste(c(min,',',max),sep = "",collapse = "")
  ) 

# transpose mean
rst <- mean %>% 
  group_by(visitn,stra) %>% 
  pivot_longer(c("cnt","mnstd","median","q13","ext"), names_to = "para", values_to = "value") %>% 
  select(-(mean:q3)) 
```

