---
title: "Chisq Test Function"
author: "Xinlong Yang"
date: "2022/11/1"
output: 
  html_document: 
    toc: yes
---


```{r setup, include=FALSE}
library(tidyverse)
```

## Chisq Test

This is a document that records the whole process of the function:func_chisq.
The following is the full code of chisq-test.

<details>
  <summary><b>Chisq Test Function Code</b></summary>
```{r myfunction}
my_chisq <- function(x){
  # input:
  # x:a matrix
  n.row <- nrow(x)
  n.col <- ncol(x)
  df <- (n.row-1)*(n.col-1)
  # the sum of row
  r <- tibble()
  # the sum of col
  c <- tibble()
  
  for (i in 1:n.row ) {
    r[1,i] <- sum(x[i,1:n.col])
  }
  for (i in 1:n.col ) {
    c[1,i] <- sum(x[1:n.row,i])
  }
  total <- sum(x)
  
  # 2×2 tables
  if (n.row==2 & n.col==2) {
    stat <- ((abs(x[1,1]*x[2,2]-x[1,2]*x[2,1])-total/2)**2*total/(r[1,1]*r[1,2]*c[1,1]*c[1,2]))[1,1]
  }
  # 2×c tables
  else if(n.row==2 & n.col>=3) {
    oc <- tibble()
    for (j in 1:n.col){
      oc[1,j] <- x[1,j]**2/c[1,j]
    }
    stat <- ((total**2/r[1,1]/r[1,2])*(sum(oc)-r[1,1]**2/total))[1,1]
  }
  # r×c tables
  else if (n.row>=2 & n.col>=3) {
    orc <- tibble()
    for (i in 1:n.row){
      for (j in 1:n.col){
        orc[i,j] <- x[i,j]**2/r[1,i]/c[1,j]
      }
    }
    stat <- total*(sum(orc)-1)
  }

  p.value <- 1 - pchisq(stat, df=df, ncp=0, log = FALSE)
  # result <- list(stat=stat,pvalue=p.value)
  result <- c(stat,p.value)
  return(result)
}
```
</details>

## Validation Case

### Case 1

This is for 2×2 tables.

The formula of Chisq Statistics for 2×2 contingency tables:

<p style="text-align: center;">$\chi 2 = \frac{(|O_{11}O_{22}-O_{12}O_{21}|-\frac{T}{2})^{2}T}{R_{1}R_{2}C_{1}C_{2}}$</p>

The Validation Data:
```{r, echo=FALSE}
m <- as.table(rbind(c(50, 250), c(5, 195)))
dimnames(m) <- list(gender = c("F", "M"),
                    party = c("Democrat","Independent"))
m
```

The statistics and P value:

```{r, echo=FALSE}
my_chisq(m)
```

### Case 2

This is for 2×c tables(c>=3).

The formula of Chisq Statistics for 2×c contingency tables:

<p style="text-align: center;">$\chi 2 = \frac{T^{2}}{R_{1}R_{2}}[\sum (\frac{O_{1j}^{2}}{C_{j}})-\frac{R_{1}^{2}}{T}]$</p>

The Validation Data:
```{r, echo=FALSE}
m1 <- as.table(rbind(c(37,49,23), c(150,100,57)))
dimnames(m1) <- list(gender = c("F", "M"),party = c("Democrat","Independent","Others"))
m1
```

The statistics and P value:

```{r, echo=FALSE}
my_chisq(m1)
```

### Case 3

This is for r×c tables(r>+3 and c>=3).

The formula of Chisq Statistics for r×c contingency tables:

<p style="text-align: center;">$\chi 2 = T[\sum (\frac{O_{ij}^{2}}{R_{i}C_{j}}-1)]$</p>

The Validation Data:
```{r, echo=FALSE}
m2 <- as.table(rbind(c(67,9,10,5), c(32,23,20,4),c(10,11,23,5)))
dimnames(m2) <- list(gender = c("F", "M","O"),
                     party = c("Democrat","Independent","Others","None"))
m2
```

The statistics and P value:

```{r, echo=FALSE}
my_chisq(m2)
```

## Some Problems

### Data Structure of Statistics

object:(r c total) are the common objects used in three cases.

```{r,eval = FALSE}
  r <- tibble()
  c <- tibble()
  
  for (i in 1:n.row ) {
    r[1,i] <- sum(x[i,1:n.col])
  }
  for (i in 1:n.col ) {
    c[1,i] <- sum(x[1:n.row,i])
  }
  total <- sum(x)
```

the code of computing chisq statistics:

Object x is a matrix.

```{r,eval = FALSE}
# this is matrix manipulation and the result is also a matrix that can't be used directly to argument in function:pchisq
stat <- ((abs(x[1,1]*x[2,2]-x[1,2]*x[2,1])-total/2)**2*total/(r[1,1]*r[1,2]*c[1,1]*c[1,2]))[1,1]
# like above
stat <- ((total**2/r[1,1]/r[1,2])*(sum(oc)-r[1,1]**2/total))[1,1]
# this is a vector manipulation
stat <- total*(sum(orc)-1)
```

The chisq statistics entered into Function:pchisq must be a vector.

Use <b>as.vector()</b> to change a matrix to a vector.


